\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-06 paper/main.tex
mk pass

@@

Assuming that viscous forces are dominating the ventricular CSF flow,
we model the flow by the Stokes equations.
The net forces of motile cilia acting on the CSF are modeled as a steady,
bidirectional tangential traction acting on parts of the ventricular wall
(\Crefpanels{fig:fig1}{d}), with free velocity slip at the walls.
Simulations show that this traction alone
leads to a partial compartmentalization of the ventricular system,
with large-scale vortex structures in the anterior, middle and
posterior ventricles (\Crefpanels{fig:fig1}{e}). The highest flow
speeds ($26.7$ \textmu m/s) occur in the vicinity of the dorsal cilia
in the middle ventricle. To model the cardiac pulsations, we impose a
pulsatile pressure difference between the anterior and posterior
boundaries (\Crefpanels{fig:fig1}{d, f}). The resulting flow
is pulsatile and directional (\Crefpanels{fig:fig1}{g}), and
in the absence of the cilia, the vortex structures vanish. The
velocity magnitude peaks at $4.8$ \textmu m/s. CSF initially flows
rostrocaudally before changing direction to caudorostral in the middle
of the cardiac cycle.

@@

\subsection{Model parameters and sources of discrepancies}
Even though we identified similar dynamics in simulations and
in vivo experiments, the transport dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the geometry of the
ventricular geometry impacts solute distribution,
and the ventricular size varies substantially 
in the control population at 2 dpf.
Additionally, differences may result
from the definition of the photoconversion site,
since the volume of this region determines the total amount of solute
and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we considered as impermeable in our simulations.
The neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
If this is also the case at 2 dpf, Dendra2 
(which are 26 kDa in size~\cite{Gurskaya2006EngineeringLight})
would leak out of the brain ventricles over the relevant time spans
and render the dynamics slower.
It may also be the case that our model predicts faster
transport because the Dendra2 diffusion
coefficient we used (based on an experimental value
\cite{GuraSadovsky2017MeasurementExpansion}) is too high.
This could be because embryonic CSF has higher viscosity
than the solution considered in the measurements, 
or because the presence of proteins, nutrients and other solutes
obstructs diffusive transport
\cite{Bloomfield1998EffectsFluid, Einstein1905UberTeilchen, Cussler2009Diffusion:Systems}.
Finally, an important difference
is the three-dimensional nature of the computational model,
while the experimental images only show 2D slices.

@@

The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the Stokes equations
is computationally more expensive
than \fixme{conventional} methods such as a
Taylor-Hood discretization
($P_k-P_{k-1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (such as $P_1-P_1$ with a velocity stabilization)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
Resultingly $\normltwo{\nabla\cdot\uu_h}=0$ to within solver tolerances,
an important property for the stability of the advection-diffusion
equation, especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly, since the BDM elements control the normal component of the 
velocity~\cite{Brezzi1985TwoProblems}.
For elements where this is not the case (such as Taylor-Hood elements),
the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties and increases problem complexity
(see \Cref{subsec:appendixA1} for more details).

@@

Wholemount in vivo live imaging and photoconversion
experiments were performed with zebrafish larvae at the 2 dpf stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xUAS:Signal-Dendra2})
adult animals, as described by D'Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}.
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+})
or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

@@

\subsection{Computational CSF dynamics in the brain ventricles}
The beating motion of the cilia generates CSF flow with persistent rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Crefpanels{fig:fig1}{a--c}).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0 
      \quad  \text{in } \Omega,\label{eq:stokes_eqs_cont} 
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.7 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a stress $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential traction $\bsigpar = {\bf 0}$).
We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

@@

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations (\Cref{eq:stokes_eqs}),
and $D$ is the molecular diffusion coefficient of the solute. 
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gp$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

In order to impose a smooth initial condition, we initially solve the diffusion equation
\begin{equation*}
    \pdifft{c} = \Tilde{D}\nabla^2 c
\end{equation*}
with an increased diffusion coefficient $\Tilde{D}=D\times 10^{5}$ for two timesteps.
The solutions are normalized by their respective maximum values, and subsequently
multiplied with the value of \Cref{eq:photoconversion_curve} at the times $t=\Delta t$
and $t=2\Delta t$, respectively.
The resulting concentration fields are used as initial conditions for $c(\xx, 0)$ and
$c(\xx, \Delta t)$ for all $\xx\in\Omega$, where $\Delta t$
is the timestep. The two initial conditions are required because of the
two-step temporal discretization scheme employed (see~\Cref{subsec:num_approx_adv_diff} for more details.)

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling (i) the Dendra2 Fluorescent Protein (Dendra2)
used in the photoconversion, (ii) the Starmaker+Green Fluorescent Protein (STM-GFP) reported by
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and (iii) extracellular vesicles.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
CSF, and $R$ is the radius of the molecule.
For the extracellular vesicles, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with the relatively large $R=150$ nm~\cite{Moghassemi2024ExtracellularDecade}.
For STM-GFP, we estimate the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients
are $D_1$ (extracellular vesicles), $D_2$ (STM-GFP) and $D_3$ (Dendra2) as given by \Cref{tab:diff_coeff_table}.

@@

\subsection{Numerical approximation of the Stokes equations}
The Stokes equations (\Cref{eq:stokes_eqs}) are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization~\cite{butcher2016numerical}.
We set the timestep size $\Delta t = 0.023$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM)
elements for the velocity.~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
The BDM elements have degrees of freedom associated with integral moments of the
velocity component normal to the facets of elements,
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 1 009 874.
The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
Continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.
Since the divergence of the velocity function space is the pressure function space for this scheme,
the divergence-free condition of the velocity is satisfied exactly, ensuring mass conservation.~\cite{Boffi2008FiniteProblem}

\subsection{Numerical approximation of the advection-diffusion equation}\label{subsec:num_approx_adv_diff}
The advection-diffusion equation (\Cref{eq:adv_diff_strong}) is discretized
with with a second-order, symmetric interior penalty DG
method in space~\cite{Arnold1982AnElements}, chosen due to its favorable
conservation properties. At each timestep the total number of degrees of freedom 
is 1 415 120. In time, the equation is discretized with the second-order backward
difference scheme BDF2.~\cite{Volker2016}
To accurately approximate the solute concentrations even
in the presence of strong advection (high Péclet numbers),
we use an upwind scheme for the velocity when calculating the
advective flux~\cite{Patankar2018NumericalFlow}.
As for the flow equations, we use a timestep size of
$\Delta t = 0.023$ s. We simulate transport for 1900 cardiac cycles,
with a final time of simulation $T = 856$ s.
A DG penalty parameter $\alpha = 50$ was chosen to ensure stability of the method.

@@

\newpage 
%%--------------APPENDIX--------------%%
\appendix
\section{Supplementary Theory}
\subsection{Weak formulation of the Stokes equations}\label{subsec:appendixA1}
We consider the time-dependent Stokes equations
\begin{subequations}
\begin{alignat}{2}
   \rho\pdifft{\uu} - \nabla \cdot \bsig(\uu, p)
   &= \mathbf{0} \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_mom}\\
  \nabla \cdot \uu &= 0 \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_div} 
\end{alignat}
\label{eq:appendix_stokes_eqs}% 
\end{subequations}%
in the domain $\Omega$ with boundary $\Gamma$, where
$\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ is the viscous
stress tensor, in which $\beps(\uu) = \frac{1}{2}(\nabla\uu + (\nabla\uu)^T)$
is the symmetric gradient. Furthermore, $\uu\in\VV$ is the fluid velocity
and $p\in Q$ is the fluid pressure, and $\rho$ and $\mu$ are the fluid density
and dynamic viscosity, respectively. Unless otherwise stated, we let $\bsig = \bsig(\uu, p)$.

To derive weak formulations of the continuous problem defined by the Stokes
equations (\Cref{eq:appendix_stokes_eqs}), we consider the function space
\begin{align*}
    \VV &= \bigg\{\vv \ : \ \vv\in\mathbf{H}^1(\Omega),
    \vv\cdot\nn=0 \ \mathrm{on} \ \Gamma\setminus\Gp\bigg\},
\end{align*}
for the velocity and $Q=L^2(\Omega)$ for the pressure.
Here $\Gp$ is the anterior/posterior boundary where we impose a normal pressure
boundary condition for the model versions that consider fluid motion induced by
cardiac pulsations. Bold-faced characters for a space denotes a $d$--dimensional vector space,
where $d$ is the spatial dimension of $\Omega$. For example,
$\mathbf{H}^1(\Omega) = [H^1(\Omega)]^3$ if $\Omega$ is three-dimensional.

@@

Convergence rates reported in~\Cref{subsec:numerical_verification} were
approximated in the following way. Define the error $e_k$ for a given mesh
resolution $h_k$. Based on the assumption that the error scales as
$e_k \approx C{h_k}^\eta$, where $C$ is a constant independent of $h_k$,
the convergence order $\eta$ was approximated as
\begin{equation}
    \eta = \frac{\log{(e_k/e_{k+1}})}{\log{(h_k/h_{k+1}})}.
    \label{eq:convergence_order_estimate}
\end{equation}
Consider a finite element function 
\begin{equation*}
    \phi = \phi_i\psi_i(\xx)
\end{equation*}
with basis functions $\psi_i$, where $i$ are the degrees of freedom and Einstein
summation is employed.
The infinity norm of $\phi$ was approximated from the vector
of coefficients $\left\{\phi_i\right\}$, which represent point evaluations of $\phi$ as
\begin{equation*}
    \normlinf{\phi} = \max_i{\vert \phi_i\vert}.
\end{equation*}

>>>>

\newcommand{\miro}[1]{{\color{cyan}#1}}

@@

Assuming that viscous forces are dominating the ventricular CSF flow,
we model the flow by the Stokes equations.
The net forces of motile cilia acting on the CSF are modeled as a steady,
bidirectional tangential traction acting on parts of the ventricular wall
(\Crefpanels{fig:fig1}{d}), with \miro{no-penetration condition at the walls}.
Simulations show that this traction alone
leads to a partial compartmentalization of the ventricular system,
with large-scale vortex structures in the anterior, middle and
posterior ventricles (\Crefpanels{fig:fig1}{e}). The highest flow
speeds ($26.7$ \textmu m/s) occur in the vicinity of the dorsal cilia
in the middle ventricle. To model the cardiac pulsations, we impose a
pulsatile pressure difference between the anterior and posterior
boundaries (\Crefpanels{fig:fig1}{d, f}). The resulting flow
is pulsatile and directional (\Crefpanels{fig:fig1}{g}), and
in the absence of the cilia, the vortex structures vanish. The
velocity magnitude peaks at $4.8$ \textmu m/s. CSF initially flows
rostrocaudally before changing direction to caudorostral in the middle
of the cardiac cycle.

@@

\subsection{Model parameters and sources of discrepancies}
Even though we identified similar dynamics in simulations and
in vivo experiments, the transport dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the ventricular geometry impacts solute distribution,
and the ventricular size varies substantially 
in the control population at 2 dpf.
Additionally, differences may result
from the definition of the photoconversion site,
since the volume of this region determines the total amount of solute
and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we considered as impermeable in our simulations.
The neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
If this is also the case at 2 dpf, Dendra2 
(which are 26 kDa in size~\cite{Gurskaya2006EngineeringLight})
would leak out of the brain ventricles over the relevant time spans
and render the dynamics slower.
It may also be the case that our model predicts faster
transport because the Dendra2 diffusion
coefficient we used (based on an experimental value
\cite{GuraSadovsky2017MeasurementExpansion}) is too high.
This could be because embryonic CSF has higher viscosity
than the solution considered in the measurements, 
or because the presence of proteins, nutrients and other solutes
obstructs diffusive transport
\cite{Bloomfield1998EffectsFluid, Einstein1905UberTeilchen, Cussler2009Diffusion:Systems}.
Finally, an important difference
is the three-dimensional nature of the computational model,
while the experimental images only show 2D slices.

@@

The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the Stokes equations
is computationally more expensive
than \fixme{conventional methods} \miro{ the more commonly used discretizations} such as a
Taylor-Hood ($P_2-P_{1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (e.g. MINI)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
Resultingly $\normltwo{\nabla\cdot\uu_h}=0$ to within solver tolerances,
an important property for the stability of the advection-diffusion
equation, especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly, since the BDM elements control the normal component of the 
velocity~\cite{Brezzi1985TwoProblems}.
For elements where this is not the case (such as Taylor-Hood elements),
the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties \miro{$<<$what do you mean here?} and increases problem complexity
(see \Cref{subsec:appendixA1} for more details).

@@

Wholemount in vivo live imaging and photoconversion
experiments were performed with zebrafish larvae at the 2 dpf stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xUAS:Signal-Dendra2})
adult animals, as described by D'Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}. \miro{$<<$ undef ref here}
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+})
or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

@@

\subsection{Computational CSF dynamics in the brain ventricles}
The beating motion of the cilia generates CSF flow with persistent rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Crefpanels{fig:fig1}{a--c}).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t\in(0, T]$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega\times(0, T],\\
      \nabla \cdot \uu &= 0 
      \quad  \text{in } \Omega\times(0, T],\label{eq:stokes_eqs_cont} 
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.7 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a stress $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential traction $\bsigpar = {\bf 0}$).
We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

@@

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t\in(0, T]$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega\times(0, T],
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations (\Cref{eq:stokes_eqs}),
and $D$ is the molecular diffusion coefficient of the solute. 
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gp$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

In order to impose a smooth initial condition, we initially solve the diffusion equation
\begin{equation}\label{eq:cinit}
  \begin{aligned}
    &\pdifft{c} = \Tilde{D}\nabla^2 c \quad \text{in } \Omega\times(0, 2\Delta t]\\
      &c(\xx, t=0) = 0\quad\text{ in }\Omega
  \end{aligned}\
\end{equation}
with an increased diffusion coefficient $\Tilde{D}=D\times 10^{5}$ for two timesteps.
The solutions are normalized by their respective maximum values, and subsequently
multiplied with the value of \Cref{eq:photoconversion_curve} at the times $t=\Delta t$
and $t=2\Delta t$, respectively.
The resulting concentration fields are used as initial conditions for $c(\xx, 0)$ and
$c(\xx, \Delta t)$ for all $\xx\in\Omega$, where $\Delta t$
is the timestep. The two initial conditions are required because of the
two-step temporal discretization scheme employed (see~\Cref{subsec:num_approx_adv_diff} for more details.)

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling (i) the Dendra2 Fluorescent Protein (Dendra2)
used in the photoconversion, (ii) the Starmaker+Green Fluorescent Protein (STM-GFP) reported by
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and (iii) extracellular vesicles.
To estimate their diffusion coefficients, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
CSF, and $R$ is the radius of the molecule.
For the extracellular vesicles, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with the relatively large $R=150$ nm~\cite{Moghassemi2024ExtracellularDecade}.
For STM-GFP, we estimate the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients
are $D_1$ (extracellular vesicles), $D_2$ (STM-GFP) and $D_3$ (Dendra2) as given by \Cref{tab:diff_coeff_table}.

@@

\subsection{Numerical approximation of the Stokes equations}
The Stokes equations (\Cref{eq:stokes_eqs}) are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization~\cite{butcher2016numerical}.
We set the timestep size $\Delta t = 0.023$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM)
elements for the velocity.~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
The BDM elements degrees of freedom are associated with integral moments of the
velocity component normal to the facets of elements,
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 1 009 874.
The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}. \miro{In particular,
  with BDM only the normal component of the velocity is continuous.}
Continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.
Since the divergence of the velocity function space is the pressure function space for this scheme,
the divergence-free condition of the velocity is satisfied exactly, ensuring mass conservation.~\cite{Boffi2008FiniteProblem}

\subsection{Numerical approximation of the advection-diffusion equation}\label{subsec:num_approx_adv_diff}
\miro{The advection-diffusion equation (\Cref{eq:adv_diff_strong}) is discretized in
space with quadratic DG elements using symmetric interior penalty DG method~\cite{Arnold1982AnElements}.
The scheme was chosen due to its favorable mass 
conservation properties}. At each timestep the total number of degrees of freedom 
is 1 415 120. In time, the equation is discretized with the second-order backward
difference scheme BDF2.~\cite{Volker2016}
To accurately approximate the solute concentrations even
in the presence of strong advection (high Péclet numbers),
we use an upwind scheme for the velocity when calculating the
advective flux~\cite{Patankar2018NumericalFlow}.
As for the flow equations, we use a timestep size of
$\Delta t = 0.023$ s. We simulate transport for 1900 cardiac cycles,
with a final time of simulation $T = 856$ s.
A DG penalty parameter $\alpha = 50$ was chosen to ensure stability of the method.

@@

\newpage 
%%--------------APPENDIX--------------%%
\appendix
\section{Supplementary Theory}
\subsection{Weak formulation of the Stokes equations}\label{subsec:appendixA1}
We consider the time-dependent Stokes equations
\begin{subequations}
\begin{alignat}{2}
   \rho\pdifft{\uu} - \nabla \cdot \bsig(\uu, p)
   &= \mathbf{0} \quad &&\mathrm{in} \ \Omega\times(0, T], \label{eq:BDM_stokes_mom}\\
  \nabla \cdot \uu &= 0 \quad &&\mathrm{in} \ \Omega\times(0, T], \label{eq:BDM_stokes_div} 
\end{alignat}
\label{eq:appendix_stokes_eqs}% 
\end{subequations}%
in the domain $\Omega$ with boundary $\Gamma$, where
$\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ is the viscous
stress tensor, in which $\beps(\uu) = \frac{1}{2}(\nabla\uu + (\nabla\uu)^T)$
is the symmetric gradient. Furthermore, $\uu\in\VV$ is the fluid velocity
and $p\in Q$ is the fluid pressure, and $\rho$ and $\mu$ are the fluid density
and dynamic viscosity, respectively. Unless otherwise stated, we let $\bsig = \bsig(\uu, p)$.

To derive weak formulations of the continuous problem defined by the Stokes
equations (\Cref{eq:appendix_stokes_eqs}), we consider the function space\footnote{
To simplify the exposition we focus only on spatial regularity. In particular,
$\vv\in\mathbf{H}^1(\Omega)$ is understood as $\vv(\cdot, t)\in \mathbf{H}^1(\Omega)$ for all
$t>0$. Similar convention is applied also when discussing advection-diffusion equation in \Cref{subsec:adv_diff_eq_weak_form}.
}
\begin{align*}
    \VV &= \bigg\{\vv \ : \ \vv\in\mathbf{H}^1(\Omega),
    \vv\cdot\nn=0 \ \mathrm{on} \ \Gamma\setminus\Gp\bigg\},
\end{align*}
for the velocity and $Q=L^2(\Omega)$ for the pressure.
Here $\Gp$ is the anterior/posterior boundary where we impose a normal pressure
boundary condition for the model versions that consider fluid motion induced by
cardiac pulsations. Bold-faced characters for a space denotes a $d$--dimensional vector space,
where $d$ is the spatial dimension of $\Omega$. For example,
$\mathbf{H}^1(\Omega) = [H^1(\Omega)]^3$ if $\Omega$ is three-dimensional.

@@

Convergence rates reported in~\Cref{subsec:numerical_verification} were
approximated in the following way. Define the error $e_k$ for a given mesh
resolution $h_k$. Based on the assumption that the error scales as
$e_k \approx C{h_k}^\eta$, where $C$ is a constant independent of $h_k$,
the convergence order $\eta$ was approximated as
\begin{equation}
    \eta = \frac{\log{(e_k/e_{k+1}})}{\log{(h_k/h_{k+1}})}.
    \label{eq:convergence_order_estimate}
\end{equation}
\miro{
We define the approximation of the infinity norm of a finite element function
$\phi = \phi_i\psi_i(\xx)$
as 
\begin{equation*}
    \normlinf{\phi} = \max_i{\vert \phi_i\vert}.
\end{equation*}
Here $\psi_i$ are the basis functions and $\phi_i$ are the degrees of freedom defined as point envaluations}
